\pdftitle

### 2018-05-12 maxNt_v2.tex
V2

%% Author: PGL  Porta Mana
%% Created: 2015-11-04T09:20:08+0100
%% Last-Updated: 2018-05-12T15:29:50+0200
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%Report-no: ***
\documentclass{article}

@@

A possible, precise definition is as follows. We want to predict, \ie\
assign a probability to, the population activity in a time bin, given the
population activities in the other bins:
\begin{equation}
  \label{eq:general_prediction}
  \p(\text{activity at one time bin} \|
  \text{activities at the remaining bins, and other info}).
\end{equation}
We can say that the simultaneous firing of at most two neurons
\enquote{explains} the activity if only the values of the first and second
empirical moments of the activities are relevant for the probability above:
\begin{multline}
  \label{eq:informal_sufficiency}
  \p(\text{activity at one time bin} \|
  \text{activities at the other bins, and other info}) ={}\\
  \p(\text{activity at some time bin} \|
  \text{1st \amp\ 2nd empirical moments, and other info});
\end{multline}
in statistical terms, the first and second moments are \emph{sufficient
  statistics}. Other definitions are possible, of course; the definition
above has the advantage of being mathematically very precise. In fact it
determines, from just the three basic laws of the probability calculus, the
mathematical form of the probability for the activity at any number of bins:
\begin{multline}
  \label{eq:sufficiency_maxent_family}
  \p(\text{activity at bins $1,2,\dotsc$} \|
  \text{other info}) ={}\\
  \shoveright{\int
    \pf(\text{activity at bin $1$} \|\ytheta)\,
    \pf(\text{activity at bin $2$} \|\ytheta)\dotsm
 \pf(\ytheta\| \text{other info})\,
  \di\ytheta,}
  \\[2\jot]
  \text{with \enquote{$\pf(\text{activity at time bin $i$} \|\ytheta)$}
    having maximum-entropy form}.
\end{multline}
The proof that the assumption of sufficient statistics leads to the
probability expression above is the Koopman-Pitman theorem
\cite{koopman1936,pitman1936,darmois1935} and can be found in
\citetext[\sect~4.5]{bernardoetal1994_r2000} and references therein. It can
be extended to higher moments and other statistics.

>>>>

%% Author: PGL  Porta Mana
%% Created: 2015-11-04T09:20:08+0100
%% Last-Updated: 2018-05-12T16:21:17+0200
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%Report-no: ***
\documentclass{article}

@@

A possible, precise definition is as follows. We want to predict, \ie\
assign a probability to, the population activity in a time bin, given the
population activities in the other bins:
\begin{equation}
  \label{eq:general_prediction}
  \p(\text{activity at one time bin} \|
  \text{activities at the remaining bins, and other info}).
\end{equation}
We can say that the simultaneous firing of at most two neurons
\enquote{explains} the activity if only the values of the first and second
empirical moments of the activities are relevant for the probability above:
\begin{multline}
  \label{eq:informal_sufficiency}
  \p(\text{activity at one time bin} \|
  \text{activities at the other bins, and other info}) ={}\\
  \p(\text{activity at some time bin} \|
  \text{1st \amp\ 2nd empirical moments, and other info});
\end{multline}
in statistical terms, the first and second moments are \emph{sufficient
  statistics}. Other definitions are possible, of course; the definition
above has the advantage of being mathematically very precise. In fact it
determines, from just the three basic laws of the probability calculus, the
mathematical form of the probability for the activity at any number of bins:
\begin{multline}
  \label{eq:sufficiency_maxent_family}
  \p(\text{activity at bins $1,2,\dotsc$} \|
  \text{other info}) ={}\\
  \shoveright{\int
    \pf(\text{activity at bin $1$} \|\ytheta)\,
    \pf(\text{activity at bin $2$} \|\ytheta)\dotsm
 \pf(\ytheta\| \text{other info})\,
  \di\ytheta,}
  \\[2\jot]
  \text{with \enquote{$\pf(\text{activity at time bin $i$} \|\ytheta)$}
    having maximum-entropy form};
\end{multline}
the $\ytheta$ being the Lagrange multipliers of the maximum-entropy
distribution. The proof that the assumption of sufficient statistics leads
to the probability expression above is the Koopman-Pitman theorem
\cite{koopman1936,pitman1936,darmois1935} and can be found in
\textcite[\sect~4.5]{bernardoetal1994_r2000} and references therein. It can
be extended to higher moments and other statistics.

When many data are available (and unfortunately also when they aren't), the
conditional density for the Lagrange multipliers,
$\pf(\ytheta\| \text{data, other info})$ is strongly peaked at the unique
values determined by the inverse problem \cite{meadetal1984}, and the
probability for the population activity is well-approximated by a
maximum-entropy one with those Lagrange multipliers. This kind of
distribution has been used at least since the 1990s
\cite{mackay1991,martignonetal1995,bohteetal2000,amarietal2003,schneidmanetal2006,shlensetal2006,tkaciketal2006,mackeetal2009,mackeetal2009b,roudietal2009,roudietal2009b,roudietal2009c,tkaciketal2009,gerwinnetal2009,barreiroetal2010b,mackeetal2011,mackeetal2011b,ganmoretal2011,granotatedgietal2013,mackeetal2013,tkaciketal2014b,moraetal2015,shimazakietal2015}
